---
title: "HW3Stat475"
output: html_document
date: '2022-11-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
cork <- read.csv("cork.csv")
corkV <- cork[,2:5]
head(cork)
m <- colMeans(corkV)
m
boxplot(corkV)
cov.cork <- cov(corkV)
cov.cork
cor(corkV)
qqnorm(cork$X1_North)
qqnorm(cork$X2_East)
qqnorm(cork$X3_South)
qqnorm(cork$X4_West)

```


```{r}
boxplot(corkV)
```
b)The South side has a bigger range in thickness of cork values. However, the means for the corks appear to be similar. The distributions of each region appear to be right skewed by the boxplots. 
c) According to the correlation matrix, South and West have the highest corelation, while East and West have the lowest. 
```{r}
cor(corkV)
library(GGally)
ggpairs(corkV)
```

Univariate Shapiro Test 
```{r}
apply(corkV, 2, shapiro.test)
```
Multivariate Test 
```{r}
library(mvShapiroTest)
mvShapiro.Test(as.matrix(cork[,2:5]))
```

e) The p-value for the test is 0.001. We can reject the null hypothesis that the mean vector is equal to 47 for all variables.
```{r}
library(DescTools)
HotellingsT2Test(corkV,mu =c(47,47,47,47), test = "f" )
```
f) The p-value is really small. This means we can reject the null hypothesis that the means are all equal to each other.
```{r}
HotellingsT2Test(corkV, test = "f" )
```

e)
```{r}
n <- nrow(corkV)
dim(corkV)
p <- ncol(corkV)
cov.corkV<- cov(corkV)
##Bonferroni 
upper_bonf <- mean(corkV[,1])- mean(corkV[,2]) + qt(1-.05/2,n-1)*sqrt(cov.corkV[1,1]-2*cov.corkV[1,2]+cov.corkV[3,2]/n)
lower_bonf <- mean(corkV[,1])- mean(corkV[,2]) - qt(1-.05/2,n-1)*sqrt(cov.corkV[1,1]-2*cov.corkV[1,2]+cov.corkV[2,2]/n)
print(rbind(lower_bonf, upper_bonf))

upper_bonf <- mean(corkV[,1])- mean(corkV[,3]) + qt(1-.05/2,n-1)*sqrt(cov.corkV[1,1]-2*cov.corkV[1,3]+cov.corkV[3,3]/n)
lower_bonf <- mean(corkV[,1])- mean(corkV[,3]) - qt(1-.05/2,n-1)*sqrt(cov.corkV[1,1]-2*cov.corkV[1,3]+cov.corkV[3,3]/n)
print(rbind(lower_bonf, upper_bonf))

upper_bonf <- mean(corkV[,1])- mean(corkV[,4]) + qt(1-.05/2,n-1)*sqrt(cov.corkV[1,1]-2*cov.corkV[1,4]+cov.corkV[4,4]/n)
lower_bonf <- mean(corkV[,1])- mean(corkV[,4]) - qt(1-.05/2,n-1)*sqrt(cov.corkV[1,1]-2*cov.corkV[1,4]+cov.corkV[4,4]/n)
print(rbind(lower_bonf, upper_bonf))

upper_bonf <- mean(corkV[,2])- mean(corkV[,3]) + qt(1-.05/2,n-1)*sqrt(cov.corkV[2,2]-2*cov.corkV[2,3]+cov.corkV[3,3]/n)
lower_bonf <- mean(corkV[,2])- mean(corkV[,3]) - qt(1-.05/2,n-1)*sqrt(cov.corkV[2,2]-2*cov.corkV[2,3]+cov.corkV[3,3]/n)
print(rbind(lower_bonf, upper_bonf))

upper_bonf <- mean(corkV[,2])- mean(corkV[,4]) + qt(1-.05/2,n-1)*sqrt(cov.corkV[2,2]-2*cov.corkV[2,4]+cov.corkV[4,4]/n)
lower_bonf <- mean(corkV[,2])- mean(corkV[,4]) - qt(1-.05/2,n-1)*sqrt(cov.corkV[2,2]-2*cov.corkV[2,4]+cov.corkV[4,4]/n)
print(rbind(lower_bonf, upper_bonf))

upper_bonf <- mean(corkV[,3])- mean(corkV[,4]) + qt(1-.05/2,n-1)*sqrt(cov.corkV[3,3]-2*cov.corkV[3,4]+cov.corkV[4,4]/n)
lower_bonf <- mean(corkV[,3])- mean(corkV[,4]) - qt(1-.05/2,n-1)*sqrt(cov.corkV[3,3]-2*cov.corkV[3,4]+cov.corkV[4,4]/n)
print(rbind(lower_bonf, upper_bonf))

```
Problem 2
```{r}
bnotes <- read.csv("bnotes.csv")
head(bnotes)
dim(bnotes)
#a)The qq plot for length is the plot that mosts falls on the line which makes it the most normal distributed variable in the data.Left_width has a curved slope which makes it look like it is not sampled from a normal distribution. Right_width and margin_wb is also curved so the same goes for this. Margin_wt and diagonal mostly fall on the line so this means that it is assumed that it is under the distribution.
apply(bnotes[ -1], 2, shapiro.test)
#b)The p-value reported is .024. Since this is less than .05, we can say that the sample comes from a multivariate model of 6 dimensions. 
mvShapiro.Test(as.matrix(bnotes[ , 2:7]))
```

